Comparison of dimensionality reduction schemes for derivative-free global optimization algorithms

### 2018-07-10 2018/YSC/slides/slides.tex
YSC: the final version of slides

\graphicspath{ {../pictures/}{img/} }
%set pages numeration
\setbeamertemplate{footline}[frame number]
\setbeamertemplate{headline}{}
\setlength\abovecaptionskip{-1pt}

@@

\begin{frame}
  \frametitle{Properties of the reduced problem}
  After applying the Peano-type evolvent \(\varphi(y(x))\) satisfies the uniform H{\"o}lder condition:
  \begin{displaymath}
    |\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
  \end{displaymath}
  \(\varphi(y(x))\) is non-smooth and has multiple local extremums even if \(\varphi(y)\) is unimodal.
  The latter problem is caused by loss of the information about \(N\)-d neighborhood after the transformation to the \(1\)-d space.

@@

\begin{frame}
  \frametitle{Smooth evolvent and non-univalent evolvent}
      %\centerline{\includegraphics[width=0.9\textwidth]{smooth.pdf}}
      \begin{figure}[ht]
        %\vspace*{-0.5cm}
        \subfloat{\raisebox{.02\textheight}{\includegraphics[width=.35\textwidth]{smooth.pdf} }}
        \subfloat{{\includegraphics[width=.25\textwidth]{noninjective.pdf} }}
      \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Shifted and rotated evolvents}
  \begin{figure}[ht]
    %\vspace*{-0.5cm}
    \subfloat{\raisebox{.0\textheight}{\includegraphics[width=.35\textwidth]{shifted.pdf} }}
    \subfloat{{\includegraphics[width=.35\textwidth]{rotated.pdf} }}
  \end{figure}
%  \begin{columns}
%    \begin{column}{0.5\textwidth}
%
%    \end{column}
%    \begin{column}{0.5\textwidth}
%      \centerline{\includegraphics[width=0.9\textwidth]{shifted.pdf}}
%    \end{column}
%  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Rotated evolvent}
  \begin{columns}
    \begin{column}{0.5\textwidth}

    \end{column}
    \begin{column}{0.5\textwidth}
      \centerline{\includegraphics[width=0.9\textwidth]{rotated.pdf}}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Optimization method}
  Optimization method generates search sequence \(\{x_k\}\) and consists of the following steps:
  \begin{enumerate}
    \setlength{\itemindent}{.1in}
    \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
    \item[Step 2.] Compute the evolvent \(y(x)\) and the function \(\varphi(y(x))\).
    \item[Step 3.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
    \item[Step 4.] Choose an interval \((x_{t-1}, x_t)\) with the greatest characteristic and
    compute objective \(f(y(x))\) in the point chosen using the decision rule \(d\):
    \begin{displaymath}
      x^{k+1}=d(t)\in (x_{t-1}, x_t)
    \end{displaymath}
    \item[Step 5.] If \(x_{t}-x_{t-1}<\varepsilon\) stop the method.
  \end{enumerate}
  \textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
\end{frame}

@@

\begin{frame}
  \frametitle{Conclusion and future work}
    Already done:
    \begin{itemize}
      \item
    \end{itemize}
    Future work:
    \begin{itemize}
      \item
    \end{itemize}
\end{frame}

>>>>

\graphicspath{ {../pictures/}{img/} }
%set pages numeration
%\setbeamertemplate{footline}[frame number]
%\setlength\abovecaptionskip{-3pt}

\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber
}
\setbeamertemplate{headline}{}

@@

\begin{frame}
  \frametitle{Properties of the reduced problem}
  After applying the Peano-type evolvent \(\varphi(y(x))\) satisfies the uniform H{\"o}lder condition:
  \begin{displaymath}
    |\varphi(y(x_1))-\varphi(y(x_2))|\leq H{|x_1-x_2|}^{\frac{1}{N}}, x_1,x_2\in[0,1],
  \end{displaymath}
  \(\varphi(y(x))\) is non-smooth and has multiple local and \textbf{global} extremums even if \(\varphi(y)\) is unimodal.
  The latter problem is caused by loss of the information about \(N\)-d neighborhood after the transformation to the \(1\)-d space.

@@

\begin{frame}
  \frametitle{Non-univalent evolvent}
    One can try to recover all preimages of \(y\in\mathbb{R}^N\) and make optimization method aware of their existence\footnote{R.G. Strongin. Numerical Methods in Multiextremal Problems (in Russian), 1978}. This allows reducing the effect of growing amount of local minimas after dimension reduction.
    According to the theory of Peano-type curves, each \(N\)-d point could have up to \(2^N\) preimages. For large \(N\) such preimages mining would be expensive.
      \begin{figure}[ht]
        \includegraphics[width=.25\textwidth]{noninjective.pdf}
      \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Shifted and rotated evolvents}
  To create a fixed amount of preimages one can use a pre-defined set of different evolvents. These evolvents could be shifted or rotated versions of the original one. Set of shifted evolvents\footnote{Strongin, R.G., Gergel, V.P., Barkalov, K.A. Parallel methods for global optimization problem
solving (in Russian), 2009} is theoretically proven to generate at least one pair of close preimages if images are close and it perform better than the set of rotated curves.
  \begin{figure}[ht]
    %\vspace*{-0.5cm}
    \subfloat{\raisebox{.0\textheight}{\includegraphics[width=.35\textwidth]{shifted.pdf} }}
    \subfloat{{\includegraphics[width=.35\textwidth]{rotated.pdf} }}
  \end{figure}
%  \begin{columns}
%    \begin{column}{0.5\textwidth}
%
%    \end{column}
%    \begin{column}{0.5\textwidth}
%      \centerline{\includegraphics[width=0.9\textwidth]{shifted.pdf}}
%    \end{column}
%  \end{columns}
\end{frame}

\begin{frame}
  \frametitle{Smooth evolvent}
  Smooth functions are more predictable for optimizer, so smooth approximation of the Peano-like \(y(x)\) curve could impove convergence rate \footnote{Goryachih, A. A class of smooth modification of space-filling curves for global optimization problems, NET 2016}.
  \begin{figure}[ht]
    \includegraphics[width=0.45\textwidth]{smooth.pdf}
  \end{figure}

\end{frame}

\begin{frame}
  \frametitle{Optimization method}
  Optimization method generates search sequence \(\{x_k\}\) and consists of the following steps:
  \begin{enumerate}
    \setlength{\itemindent}{.1in}
    \item[Step 1.] Sort the search information (one-dimensional points) in increasing order.
    \item[Step 2.] Compute the evolvent \(y(x^k)\) and the function \(\varphi(y(x^k))\).
    \item[Step 3.] For each interval \((x_{i-1}, x_i)\) compute quantity \(R(i)\), called characteristic.
    \item[Step 4.] Choose an interval \((x_{t-1}, x_t)\) with the greatest characteristic and
    compute objective \(f(y(x))\) in the point chosen using the decision rule \(d\):
    \begin{displaymath}
      x^{k+1}=d(t)\in (x_{t-1}, x_t)
    \end{displaymath}
    \item[Step 5.] If \(x_{t}-x_{t-1}<\varepsilon\) stop the method.
  \end{enumerate}
  \textit{\footnotesize	{Detailed description: Strongin R.G., Sergeyev Ya.D.: Global optimization with non-convex constraints. Sequential and parallel algorithms (2000), Chapter 7}}
\end{frame}

@@

\begin{frame}
  \frametitle{Conclusions}
    %Already done:
    \begin{itemize}
      \item the smooth evolvent and the non-univalent one demonstrate the best result in the problems of small dimensionality and can be applied successfully in solving the problems with the computational costly objective functions.
      \item the shifted evolvents introduce large overhead costs on the execution of the method due to the requirement to adding an auxiliary constraint. About 95\% of iterations are overhead to fight the auxiliary constraint.
      \item rotated evolvents perform almost the same as the shifted ones but without overhead.
    \end{itemize}
    %Future work:
    %\begin{itemize}
    %  \item
    %\end{itemize}
\end{frame}
